---
title: "Diamonds Dataset Visualization"
author: "By Siriwut Thongmak"
date: "2023-02-25"
output:
  html_document: default
  pdf_document: default
---

### What's am I doing?

Explore Dataset `diamonds` and create 5 visualizations.

### Explore the data
```{r message = FALSE, warning=FALSE}
library(tidyverse)
head(diamonds)
```

Dataset Description:

1. `carat`: weight of the diamond (0.2--5.01)
2. `cut`: quality of the cut (Fair, Good, Very Good, Premium, Ideal)
3. `color`: diamond color, from J (worst) to D (best)
4. `clarity`: a measure of the number and size of inclusions within the diamond, from I1 (worst), SI2,                  SI1, VS2, VS1, VVS2, VVS1, IF (best)
5. `depth`: total depth percentage = z / mean(x, y) = 2 * z / (x + y) (43--79)
6. `table`: width of top of diamond relative to widest point (43--95)

There are x, y, z data, so let's calculate the volume

### Transforming Data

```{r message = FALSE, warning=FALSE}
library(tidyverse)
diamonds <- diamonds %>%
  mutate(volume = x * y * z) %>%
  select(-c(x, y, z))
head(diamonds)
```


### Chart 1 - carat, dept, clarity and color Vs price
```{r chart1, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
library(patchwork)
p1 <- ggplot(diamonds %>% sample_frac(0.1),aes(carat,price, col=cut))+
        geom_point(size = 1, alpha=0.5) +
        geom_rug() +
        theme_minimal() +
        labs(title = "Relationship between carat and price",
        x = "Carat",
        y = "Price USD",
        caption = "Datasource: Diamonds") +
        scale_color_viridis_d(direction = 1)

p2 <- ggplot(diamonds %>% sample_frac(0.1),aes(depth,price, col=cut))+
        geom_point(size = 1, alpha=0.5) +
        theme_minimal() +
        labs(title = "Relationship between Depth and price",
        x = "Depth",
        y = "Price USD",
        caption = "Datasource: Diamonds") +
        scale_color_viridis_d(direction = 1)

p3 <- ggplot(diamonds %>% sample_frac(0.1),aes(clarity,price, col=cut))+
        geom_col() +
        theme_minimal() +
        labs(title = "Relationship between Clarity and Price",
        x = "Clarity",
        y = "Price USD",
        caption = "Datasource: Diamonds") +
        scale_color_viridis_d(direction = 1)

p4 <- ggplot(diamonds %>% sample_frac(0.1),aes(color,price, col=cut))+
        geom_col() +
        theme_minimal() +
        labs(title = "Relationship between Color and Price",
        x = "Color",
        y = "Price USD",
        caption = "Datasource: Diamonds") +
        scale_color_viridis_d(direction = 1)

(p1+p2)/p3/p4
```



---------------------------------------------------------------------------------------------

### Chart 2 - Count of Cut type

```{r chart2, message=FALSE, warning=FALSE}
ggplot(diamonds, aes(cut, fill = cut))+
  geom_bar(position = "dodge")+
  theme_minimal() +
  labs(title = "Count of Cut type",
       x = "Cut",
       y = "Count",
       caption = "Datasource: Diamonds")
```



---------------------------------------------------------------------------------------------

### Chart 3 - Density plot of price by color
```{r chart3, message=FALSE, warning=FALSE}
ggplot(diamonds, aes(price, fill = color))+
  geom_density(size = 0, alpha = 0.5, adjust = 2)+
  theme_minimal() +
  labs(title = "Density plot of price by color",
       x = "price",
       caption = "Datasource: Diamonds")
```



---------------------------------------------------------------------------------------------

### Chart 4 - carat vs volume

```{r chart4, message=FALSE, warning=FALSE}
set.seed(42)
ggplot(diamonds %>% sample_n(5000), aes(carat,volume, col = cut))+
  geom_point(size = 3, alpha = 0.3)+
  geom_rug()+
  theme_minimal() +
  labs(title = "carat vs volume",
       x = "carat",
       caption = "Datasource: Diamonds") +
  scale_color_viridis_d(direction = 1)
```



---------------------------------------------------------------------------------------------

### Chart 5 - Count of Diamonds by Color and Clarity

```{r chart5, message=FALSE, warning=FALSE}
ggplot(diamonds, aes(x = clarity, y = color, col = ..n..)) +
  geom_count() +
  theme_minimal()+
  scale_color_gradient(low = "gold",high="purple")+
  labs(title = "Count of Diamonds by Color and Clarity",
       x = "Clarity",
       y = "Color",
       caption = "Datasource: Diamonds")
```


